---
title: "a1_task1_chiong_genevieve"
author: "Genevieve Chiong"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(patchwork)
library(readxl)
library(lubridate)
library(scales)
```


```{r}
# Read in the Sierra Amphibians data 
sierra_amphibians <- read_excel(here("data", "sierra_amphibians.xlsx"))
```

```{r}
# Wrangle data to convert the date to a date and pull the year, filter out to exclude the 'EggMass' life stage level, group by year, and count the total mountain yellow-legged frog (Rana muscosa) observations for each year across all water bodies 

ramu_counts <- sierra_amphibians %>% 
  mutate(date = ymd(survey_date)) %>%  
  mutate(year = year(date)) %>%  
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage %in% c("Adult", "SubAdult", "Tadpole")) %>% #
  group_by(year) %>% 
  count(amphibian_life_stage, wt = amphibian_number)

# Create a graph of total mountain-yellow-frog counts each year across all water bodies, by life stage
ggplot(data = ramu_counts, 
       aes(x = year, 
           y = n)) +
  geom_line(aes(color = amphibian_life_stage)) +
  scale_x_continuous(breaks = seq(from = 1995,
                                  to = 2002,
                                  by = 1)) +
  scale_y_continuous(labels = comma,
                     breaks = seq(from = 0,
                                  to = 175000,
                                  by = 25000)) + 
  labs(x = "\nYear",
       y = "Mountain Yellow-Legged Frog Counts\n")
  
```

```{r}
# Wrangle data to count the observed combined adult and subadult yellow-legged frogs (Rana muscosa) in the 5 lakes with the greatest total observed counts 

adult_subadult_ramu <- sierra_amphibians %>% 
  filter(amphibian_species == "RAMU") %>% 
  filter(amphibian_life_stage %in% c("Adult", "SubAdult")) %>% 
  group_by(lake_id) %>% 
  count(wt = amphibian_number) %>% 
  slice_max(n, n = 5)

```
    
    
    
    
```{r}
# Create a column graph displaying Lake ID vs total yellow-legged frog counts (adult + subadult) for the 5 lakes 

ggplot(data = adult_subadult_ramu,
       aes(x = lake_id,
       y = n))
```


     count(lake_id, wt = amphibian_number) 
   
   fct_reorder(lake_id, amphibian_number) %>% 
  slice_max(lake_id, n=5)

mutate(month_name = fct_reorder(month_name, record_month))


 
 fct_reorder()
  slice_max(lake_id, n = 5)

x total counts 

y = lake id? 

 group_by(lake_id) %>% 
  count(wt = amphibian_number) %>% 
  slice_max(lake_id, n = 5)


  top_5_status <- sf_trees %>% 
  count(legal_status)  %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5)
  
Make a graph of those top 5 observations by legal status. 

ggplot(data = top_5_status,
       aes(x = fct_reorder(legal_status, tree_count),
       y = tree_count)) +
  geom_col() + 
  labs(x = "Legal Status",
       y = "Tree Count") +
  coord_flip() + 
  theme_minimal()
  
  
```{r}

```


### Citations 
Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3













